---
title: "bayesian"
author: "Amr Attyah"
date: "12/31/2018"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

*Management at a call center is investigating the call load in order to find an efficient staffing policy. Assume that time intervals between calls are exponentially 
distributed. Assume the mean time between calls is constant during the mid-morning period. The following sequence of call times was collected during mid-morning, measured in seconds after the start of data collection: 168, 314, 560, 754, 1215, 1493, 1757, 1820, 1871,1982, 2134, 2430, 3187, 3388, 3485. Assume an inverse Gamma prior distribution with shape a =4 and scale b = 0.0015 for the mean time in seconds between calls Q. Find the posterior distribution for Q. Find the prior and posterior mean and standard deviation for Q. Discuss. (Note: Because of the memoryless property of the exponential distribution, you can treat the time until the first call as having an exponential distribution.)*

&nbsp;

a1=shape=a0+n=4+15 = 19
b1=scale=1/(b0^(-1) +sum(x_i)) = 1/(0.0015^(-1)+3485 seconds) = 0.000241

```{r}
library('invgamma')


CallsBySec <- c(168, 314, 560, 754, 1215, 1493, 1757, 1820, 1871, 
                1982, 2134, 2430, 3187, 3388, 3485)


alpha0 = 4        # Prior shape
beta0 = 0.0015      # Prior inverse-scale (scale of the Gamma distributionf for birth rate)

n = length(CallsBySec) # number of calls
sum.ibt =CallsBySec[n] 
alpha1 = alpha0 + n   # posterior shape
beta1 = (1/beta0 + sum.ibt)^-1

thetavals=seq(length=100,from=0.1,to=600)
prior.dens=dinvgamma(thetavals,alpha0,scale=beta0)
norm.lik=dinvgamma(thetavals,n-1,scale=1/sum.ibt)
post.dens=dinvgamma(thetavals,alpha1,scale=beta1)
plot(thetavals,prior.dens,type="l",col="blue",main="Triplot for Mean Time Between Calls",
     xlab="Mean Time Between Calls (Seconds)",ylab="Probability Density",
     xlim=c(0,600),ylim=c(0,0.01))
lines(thetavals,norm.lik,col="green")
lines(thetavals,post.dens,col="red")
legend(0.01,5.9,c("Prior","Norm Lik","Posterior"),col=c("blue","green","red"),lty=c(1,1,1))
```

Starting with a very small beta resulted in a very spread out  inverse gamma distribution with mean 222.22 and a large standard deviation of 157.1. Using the exponential data, which is the conjugate pair of the inverse gamma, we updated our beliefs to have a mean of 230 and standard deviation of 55.91. The graph displayed about confirms the findings that the posterior distribution is narrower and and more concentrated.

## Problem 2

*Use the results of Problem 1 to find the posterior distribution
for the mean number of calls per second at the call center during mid-morning. Find 95% credible intervals for the mean number of calls per second and for the mean time between calls. 
Explain your reasoning.*

&nbsp;

If the mean time between calls has a inverse gamma distribution (shape = $\widehat{\alpha}$,scale = $\widehat{\beta}$), then the mean number of calls has a gamma distribution(shape = $\widehat{\alpha}$,scale = $\widehat{\beta}$), where $\theta$ = $\lambda$^-(1)

```{r}
lambda =  seq(length=100,from=0.001,to=0.02)
prior.dens=dgamma(lambda,shape= alpha0,scale =beta0)
post.dens=dgamma(lambda,shape = alpha1,scale =beta1)
plot(lambda,prior.dens,type="l",col="blue",main="Gamma Posterior & Prior Distributions",
     xlab="Call Rate (Per Second)",ylab="Probability Density",
     xlim=c(0,0.02),ylim=c(0,500))
lines(lambda,post.dens,col="red")
legend(0.01,400,c("Prior","Posterior"),col=c("blue","red"),lty=c(1,1,1))

cat("95% credible intervals for the mean number of calls",qgamma(c(0.025, 0.975),shape = alpha1,scale =beta1))
cat("95% credible intervals for the mean time between calls",qinvgamma(c(0.025, 0.975),shape = alpha1,scale = beta1))

```



Prior Mean = a0 $\times$ b0 = 4 $\times$ 0.0015 = 0.006 

Prior Standard Deviation = sqrt(a0 $\times$ b0^2) = sqrt(4 $\times$ 0.00152^2)=0.003

Posterior Mean = a1 $\times$ b1 = 19 $\times$ 0.000241 = 0.004579

Posterior Standard Deviation = sqrt(a1 $\times$ b1^2)= sqrt(19 $\times$ 0.000241^2) = 0.00105
